Using numerical modelling to investigate temperature and wave exposure as drivers

### 2019-10-09 morphology_paper.Rmd
moved a sentence from results FB wave spectra to discussion. going to test a word output now.

To clarify the averaged wave exposure maps presented, the propagation of a typical offshore wave spectrum as produced from a single time-step in SWAN is presented in figure \ref{fig:FB spectrum}. Tracing the wave height contours into False Bay its clear why this bay’s western periphery is predominantly sheltered. It should be mentioned that some of the annual winter frontal depression systems pass the Cape Peninsula from the west to east, resulting in wave propagating towards the continent from much more southerly directions. This results in positive and negative wave exposure anomalies all around the peninsula.

@@

There were clear patterns and clines in both the temperature and wave exposure data. The thermal and wave exposure regimes around the Cape Peninsula are driven by a complex interaction between wind, temperature and wave metrics, which do not act independently but instead influence one another. The hydrodynamic regime is modulated by the relative sheltering against the predominant south-westerly swell direction and the size of the fetch for local wave generation. The direction of the dominant swell changes to the south west in winter, generated by strong low pressures that originate from the southern ocean [@Reason2006], which False Bay is sheltered from [@Shipley1964; @Atkins1970-bx; @Dufois2012-vy]. In summer, these swells rotate anticlockwise and are able to enter False Bay, providing an increased variability of Hs and Tp in this region. It should be noted that what is classified as sheltered around the South African coastline (a high energy coastline) might be classified as exposed in other regions of the world [@Norderhaug2012-ks; @Leliaert2000a]. It is due to the near consistent south-westerly swell and the complex orography around the peninsula that the wave energy distribution around the Cape Peninsula varies significantly over a small geographical area.
The directional sheltering effect of the Cape Peninsula, against the dominant swell direction is clearly observed in the wave exposures presented in figure \ref{fig:wave power map}. Increased wind speed at sites along the west side of the peninsula in a southerly direction trigger upwelling events along the western side of the peninsula [@Rouault2010-gi; @Field1980a]. In general, the western side of the peninsula is more exposed than the eastern side (False Bay) and experiences significanlty higher significant swell height and significant peak period. This was also reflected in the wave exposure categories which show a trend of decreasing wave exposure (wave power) around the peninsula and into False Bay. There are also differences between the types of waves that each side of the peninsula experiences. The wave analysis shows the western side of the peninsula is not only more exposed but the hydrodynamics are largely swell driven, while on the eastern side most of the wave energy is driven by a combination of swell and wind with the latter being the predominant driver.

>>>>

To clarify the averaged wave exposure maps presented in figure \ref{fig:wave power map}, the propagation of a typical offshore wave spectrum as produced from a single time-step in SWAN is presented in figure \ref{fig:FB spectrum}. Tracing the wave height contours into False Bay its clear why this bay’s western periphery is predominantly sheltered.

@@

There were clear patterns and clines in both the temperature and wave exposure data. The thermal and wave exposure regimes around the Cape Peninsula are driven by a complex interaction between wind, temperature and wave metrics, which do not act independently but instead influence one another. The hydrodynamic regime is modulated by the relative sheltering against the predominant south-westerly swell direction and the size of the fetch for local wave generation. The direction of the dominant swell changes to the south west in winter, generated by strong low pressures that originate from the southern ocean [@Reason2006], which False Bay is sheltered from [@Shipley1964; @Atkins1970-bx; @Dufois2012-vy]. In summer, these swells rotate anticlockwise and are able to enter False Bay, providing an increased variability of Hs and Tp in this region. It should be noted that what is classified as sheltered around the South African coastline (a high energy coastline) might be classified as exposed in other regions of the world [@Norderhaug2012-ks; @Leliaert2000a]. It is due to the near consistent south-westerly swell and the complex orography around the peninsula that the wave energy distribution around the Cape Peninsula varies significantly over a small geographical area. The directional sheltering effect of the Cape Peninsula, against the dominant swell direction is clearly observed in the wave exposures presented in figure \ref{fig:wave power map}.It should be mentioned that some of the annual winter frontal depression systems pass the Cape Peninsula from the west to east, resulting in wave propagating towards the continent from much more southerly directions. This results in positive and negative wave exposure anomalies all around the peninsula. Increased wind speed at sites along the west side of the peninsula in a southerly direction trigger upwelling events along the western side of the peninsula [@Rouault2010-gi; @Field1980a]. In general, the western side of the peninsula is more exposed than the eastern side (False Bay) and experiences significanlty higher significant swell height and significant peak period. This was also reflected in the wave exposure categories which show a trend of decreasing wave exposure (wave power) around the peninsula and into False Bay. There are also differences between the types of waves that each side of the peninsula experiences. The wave analysis shows the western side of the peninsula is not only more exposed but the hydrodynamics are largely swell driven, while on the eastern side most of the wave energy is driven by a combination of swell and wind with the latter being the predominant driver.
